%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                        A Tex File Made by Gau Syu
%                              GauSyu@Gmail.com
%
%###########################################
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                     Include The Common Settings
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Common}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                                Letters and Notations
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Letters_and_Notations}
\loadglsentries{Glossaries/TableOfCats}
\loadglsentries{Glossaries/TableOfSymbols}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                                 PDF File Information
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypersetup{
             pdftitle={Seminar notes on Algebra},
             pdfauthor={Gau Syu},
             pdfsubject={algebra},
             pdfkeywords={seminar, notes},
             pdfproducer={XeLaTeX},
    colorlinks=true,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                                           Title
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{{\giant Seminar notes on}\\ \Giant\textsc{Algebra}}
\author{Gau Syu}
\date{
  \begin{gather*}
    \xymatrix@=5em{
    & Y' \ar@{-->}[dr]^g \ar[ddl]^(0.65){[1]} & \\
    Z' \ar@{-->}[ur]^f \ar[d]_{[1]} & &
        X' \ar[ll]_{j[1] \circ i}^{[1]}|(0.25)\hole|(0.75)\hole
           \ar[ddl]_(0.4)i^(0.4){[1]}|(0.25)\hole|(0.5)\hole \\
    X \ar[rr]^{v \circ u} \ar[dr]_u & &
        Z \ar[u] \ar[uul]\\
   & Y \ar[ur]_v \ar[uul]_(0.6)j|(0.5)\hole|(0.75)\hole &
    }\\
    \\
    \\
  \end{gather*}
\emph{\small Last Update: \today}}

\begin{document}
\frontmatter
\maketitle
\dominitoc
\include{Preface}
\tableofcontents
\mainmatter
\pagestyle{plain}
\setcounter{minitocdepth}{2}
\part{Mathematical Structures}
  The notion of ``Mathematical structure'', as an explicit mathematical concept, first appears in N.Bourbaki's \emph{Elements of Mathematics}, Volume 1 \emph{Theory of Sets}, Chapter 4 ``Structures''. By using axiomatic set theory, Bourbaki suggested a general definition of ``mathematical structure'' and ``isomorphism'', although this theory was rarely used even in his own book.

  Category Theory, as newborn fundamental theory, provides a more suitable language to describe mathematical structures and influence modern mathematics and other subjects profoundly. Hence more and more modern algebra textbooks involve an introduction to category theory.

  In this part, we introduce some knowledge of category theory we will mention in the rest of this book. Most of them can be found in every textbook of category theory.

  The standard reference is MacLane's famous \emph{Categories for the Working Mathematician} \cite{lane1998categories}.
  In addition, F.W.Lawvere and S.H.Schanuel's \emph{Conceptual Mathematics} \cite{lawvere1997conceptual}, S.Awodey's \emph{Category Theory} \cite{awodey2010category} are straightaway textbooks.
  F.Borceux \emph{Handbook of categorical algebra} \cite{borceux} is an informative Encyclopedia.

  Although unnecessary in principle, having basic algebraic training is helpful to understand this part. One can refer
  R.Godement's \emph{Cours d'alg{\`e}bre} \cite{godement1963cours}, N.Jacobson's \emph{Basic Algebra} \cite{jacobson1980basic}, M.Artin's \emph{Algebra} \cite{artin2011algebra} and even any undergraduate textbook of algebra.

  List of chapters:

  The Language of Categories

  Set Theory

  Theory of Limits

  Special Objects and Morphisms

  Universal Structures

  Internal Category Theory

  \include{MathStructure/Category}
  \include{MathStructure/Limits}
  \include{MathStructure/Objects_and_morphisms}
  \include{MathStructure/Universal}

%concrete category


  \include{MathStructure/Order}



\part{Group Theory}
%Lattice and monoids


\part{Commutative Objects}
\chapter{Abelian Groups}
\chapter{Abelian Category}
\chapter{Commutative Algebra}
\chapter{Algebraic Number Theory}
\chapter{Algebraic Geometry}
\chapter{Arithmetic Geometry}

\part{Associative Objects}
\chapter{Groups in Category}

\chapter{Associative Algebra}

\chapter{Field Extension}


\part{Galois Objects}
\chapter{Galois Theory}

\part{Lie Objects}

\part{Homology Objects}

%\part{Weird Objects}

%\part{Galois Theory}
%\include{Chapter5}%Algebraic Extensions
%\include{Chapter6}%Galois Theory
%infinite Galois Theory

\part{Dedekind Domain}
\section{23}


\begin{appendices}
%\renewcommand{\thepart}{}%\Alph{part}.}
\CTEXsetup[name={,},number={}]{part}
\part{Appendix}
%\renewcommand{\thechapter}{}
%\include{CategoryTheory}
%\chapter{Category}

  \include{Appendix/Set}

  \include{Appendix/Books}

\end{appendices}

\newpage
\backmatter

%Chapter=Table of Categories
\printglossaries

\newpage
%Chapter=Index
\phantomsection
\addcontentsline{toc}{chapter}{Index}
\printindex

\newpage
%Chapter=Bibliography
\phantomsection
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{apalike}
\bibliography{bib}

\end{document}
